Air cargo load and route planning in pickup and delivery operations

### 2024-01-28 latex/eswa.tex
ga_deap_mmkp not working propperly
GA DEAP for ACLPP not working
Latex files updated
More comments included in the responses

\usepackage{natbib}
\bibliographystyle{elsarticle-harv}

@@

Recently, \citep{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP), and {\it Weight and Balance Problem} (WBP). Several aspects were considered: item characteristics to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs); when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how total cargo weight is balanced; etc.

@@

\begin{table}[htb]

@@

\begin{table}[h]
	\centering
	\caption{Air cargo transport: literature, sub-problems and features}  \label{tab:sa}
	\footnotesize
	\begin{tabular}{r|cccc|ccccc}
		\toprule
		Work & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\citep{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\citep{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\citep{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\citep{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\citep{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{Paquay2016,Paquay2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\citep{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\citep{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		\citep{zhao2023} & $\bigstar$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		\citep{MiguelOptimalAP} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		{\bf This article}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

@@

\begin{table}[h]
	\centering
	\caption{Aircraft parameters}  \label{tab:larger}
	\footnotesize
	\begin{tabular}{c | c c c c c c c c c}
		\toprule
		& \multicolumn{3}{c}{$Payload$: 75,000kg} & \multicolumn{3}{c}{$limit^{CG}_{long}$: $1.170m$} &
		\multicolumn{3}{c}{$limit^{CG}_{lat}$: $0.19m$} \\
		\midrule
		\multirow{2}{*}{{\boldm{$p_i$}}}  & $p_{17}$ & $p_{15}$ & $p_{13}$ & $p_{11}$ & $p_{9}$ & $p_{7}$ & $p_{5}$ & $p_{3}$ & $p_{1}$ \\
		& $p_{18}$ & $p_{16}$ & $p_{14}$ & $p_{12}$ & $p_{10}$ & $p_{8}$ & $p_{6}$ & $p_{4}$ & $p_{2}$ \\
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{long}$} ($m$)} & -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\
		& -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\			
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{lat}$} ($m$)}  & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 \\
		& -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 \\	
		\midrule
		{\boldm{$W_i$}} ($kg$)      &   4,500   &    4,500  &   4,500   &  4,500    & 4,500     & 4,500     & 4,500     & 3,000    & 3,000   \\
		{\boldm{$V_i$}} ($m^3$)   &   14.8   &   14.8   &  14.8    &  14.8    & 14.8     & 14.8     & 14.8     & 10.0    & 7.0 \\	
		\midrule	
		\textbf{Fuel cost}  & \multicolumn{9}{c}{ $c_d$ = US\$ $4.90/km$ }	 \\
		\midrule
		\textbf{Fuel consumption rate}  & \multicolumn{9}{c}{ $c_g = 5\%$} \\		
		\midrule
		\textbf{Maximum weight}  & \multicolumn{9}{c}{ $W_{max} = \sum_i W_i = 75,000kg$} \\
		\bottomrule
	\end{tabular}
	\normalsize 
\end{table}

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.22]{lateral.png}
	\caption{Aircraft longitudinal cut, where red lines are pallets positions}
	\label{fig:lateral}
\end{figure}

@@

\begin{table}[ht]
	\centering
	\footnotesize
	\caption{Problem structure}  \label{tab:struc}
	\begin{tabular}{l l}
		\toprule
		\textbf{Notation}	& \textbf{Description} \\
		\midrule
		$L = \{ 0, 1, \ldots, K \}$ & Set of $K+1$\/ nodes of the tour, where node 0 is the base \\
		\midrule
		$\pi$               & A permutation between nodes 1, \ldots, $K$\\
		\midrule
		$S_K$               & Set of $K!$\/ permutations \\
		\midrule
		$\pi_k$             & The $k^{th}$\/ node of tour $\pi$, $1 \leq k \leq K$\\
		\midrule
		Tour $\pi$          & $\{0,\pi_1,\ldots, \pi_K, 0\}$ \\
		& For ease of notation, $\pi_0 = \pi_{K+1} = 0$\\
		\midrule
		$L_k$               & Set of remaining nodes of tour at node $k$, $0 \leq k \leq K$\\
		& By definition, $L_0 = L$\\
		\midrule
		$d(a,b)$            & Distance from node $a$\/ to node $b$, where $0 \leq a, b \leq K$ \\
		&   By definition, $d(a,a) = 0, \forall a$\\
		\midrule
		$C=\left[c_{a,b}\right]$    & Cost matrix of flights, where $c_{a,b} = c_d \times d(a,b)$\\
		\midrule
		$M = \{1, \ldots, m \}$  & Set of $m$\/ pallets in specific positions within the aircraft \\
		&  See Table \ref{tab:larger}, where $m=18$ \\
		\midrule
		$N_k = \{1, \ldots, n_k \}$ & Set of $n_k$\/ items available for loading at node $k$, $1 \leq j \leq n_k$, $0 \leq k \leq K$\\
		\midrule
		$N = \bigcup_{0 \leq k \leq K} N_k$ & Set of items in all nodes along a tour\\
		\midrule
		$Q_k = \{1, \ldots, m_k \}$ & Set of $m_k \leq m$\/ packed contents at node $k$, $1 \leq q \leq m_k$, $0 \leq k \leq K$ \\ 
		& By definition, $m_0 = 0$ and $Q_0 = \varnothing$\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[H]
	\centering
	\footnotesize
	\caption{Decision variables and allocation graph}  \label{tab:graph}
	\begin{tabular}{l l}
		\toprule
		\textbf{Notation}	& \textbf{Description} \\
		\midrule
		$X_{ij}^{\pi_k}$\/ and $Y_{iq}^{\pi_k}$ & Binary variables, where $1 \leq i \leq m$, $1 \leq j \leq n_{\pi_k}$, $1 \leq q \leq m_{\pi_k}$\/ and $0 \leq k \leq K$\\
		\midrule
		$X_{ij}^{\pi_k} = 1$ & If item $j$\/ at node ${\pi_k}$\/ is assigned to pallet $i$, and 0 otherwise\\
		\midrule
		$Y_{iq}^{\pi_k} = 1$ & If packed content $q$\/ at node ${\pi_k}$\/ is assigned to pallet $i$, and 0 otherwise\\
		\midrule
		$T_i^{\pi_k} \in L_{\pi_k}$ &  Destination of items and packed contents assigned to pallet $i$\ at node ${\pi_k}$\\
		\midrule
		$G_{\pi_k}(V_{\pi_k}, E_{\pi_k})$ & Allocation graph at node ${\pi_k}$\\
		\midrule
		$V_{\pi_k} = M \cup N_{\pi_k} \cup Q_{\pi_k}$ & Allocation graph vertices at node ${\pi_k}$: pallets, items and packet contents\\
		\midrule
		$E_{N_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$, corresponding to shipped items\\
		\midrule
		$E_{Q_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$, corresponding to packed contents\\
		\midrule
		$E_{\pi_k} = E_{N_{\pi_k}} \cup E_{Q_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$\\
		\midrule
		$(i, j) \in E_{N_{\pi_k}}$ & If $X_{ij}^{\pi_k} = 1$, where $i$\/ is a pallet and $j$\/ is a item at node ${\pi_k}$ \\
		\midrule
		$(i, q) \in E_{Q_{\pi_k}}$ & If $Y_{iq}^{\pi_k} = 1$, where $i$\/ is a pallet and $q$\/ is a packed content at node ${\pi_k}$\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[H]
	\centering
	\footnotesize
	\caption{Calculus functions} \label{tab:functions}
	\begin{tabular}{ c l }
		\toprule
		\textbf{Function} & \textbf{Description} \\
		\midrule
		(\ref{eq:scores}) & Total score of transported items throughout tour $\pi$\\
		\midrule
		(\ref{eq:tau})    & Longitudinal torque applied by loaded pallets at node ${\pi_k}$\\
		\midrule
		(\ref{eq:costs})  & Total cost of fuel on tour $\pi$\/ (distances and CG longitudinal deviations) \\
		\midrule
		(\ref{eq:departing}) & Set of not visited nodes at node ${\pi_k}$\\
		\midrule
		(\ref{eq:LatItem}) & Lateral torque at node ${\pi_k}$ (shipped items) \\
		\midrule
		(\ref{eq:LatPacked}) & Lateral torque at node ${\pi_k}$ (packed contents)\\
		\midrule
		(\ref{eq1}) & Objective function of ACLP+RPDP\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[H]
	\centering
	\footnotesize
	\caption{Constraints} \label{tab:constraints}
	\begin{tabular}{ c l }
		\toprule
		\textbf{Constraint} & \textbf{Description} \\
		\midrule
		(\ref{eq:torqlong}, \ref{eq:torqlat}) & Longitudinal and lateral torques must be within aircraft limits \\
		\midrule
		(\ref{eq:app2}, \ref{eq:app3}) & Items allocated to each pallet cannot exceed its weight and volume limits\\
		\midrule
		(\ref{eq:app4}) & At most, each item is associated with a single pallet \\
		\midrule
		(\ref{eq:app5}) & Packed contents that have not yet reached their destination must remain on board\\
		\midrule
		(\ref{eq22}, \ref{eq23}) & Items allocated on the same pallet must have the same destinations\\
		\midrule
		(\ref{eq24}, \ref{eq25}) & If there is a packed content on the pallet, it must also have the same destination as other items\\

@@

\begin{algorithm}[H]

@@

In the generation of test instances, we use two types of random selections:
\begin{itemize}
	\setlength\itemsep{-0.3em}
	\item $RandomInt(i_1,i_2)$: randomly selects a integer number in $[i_1,i_2]$, where $i_1$ and $i_2$ are integer numbers;
	\item $Roulette(set)$ biased through $\phi$: selects an element from $set$, where the probability of each element is proportional to the value of a given function $\phi$\/ defined on $set$.
\end{itemize}

@@

	\begin{table}[H]
		\centering
		\caption{\textit{iRace} testing results}  \label{tab:irace}
		\footnotesize
		\begin{tabular}{cccc}
			\toprule
			\textbf{surplus} & \boldm{$threshold_1$} & \boldm{$threshold_2$} & \textbf{time (min)}\\
			\midrule
			1.2& 0.8621 & 1.0539 & 47 \\
			1.5& 0.9199 & 1.1399 & 59 \\
			2.0& 0.9617 & 1.5706 & 63 \\						
			\bottomrule
		\end{tabular}
		\normalsize
	\end{table}

@@

\begin{table}[H]
{\color{blue}
	\centering
	\caption{Overall results}  \label{tab:mr}
	\footnotesize
	\begin{tabular}{cccc}	
	\toprule
	{\bf Method} & {\bf Best scenarios} & {\bf Worst scenarios} & {\bf Worst run times (min)}  \\
	\midrule
	NMO           & 4             & 5    & {\color{red}60}  \\
	ACO           & 2, 3          & 4, 5 & {\color{red}25}, {\color{red}61}  \\
	GRASP         & 1             & 4, 5 & {\color{red}28}, {\color{red}55}  \\
	TS            & -             & 5    & {\color{red}44}  \\
	\midrule
	{\it Gurobi}  & 1, 2, 3, 4    & 5    & {\color{red}did not solve}  \\
	{\it Shims}   & 1, 2, 3, 4, 5 & -    & {\bf 3.26} \\
	\bottomrule
	\end{tabular}
\normalsize
}
\end{table}

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.25]{brazil.png}
	\caption{ {\it Shims/GA} best tour (may not be optimal)}
	\label{fig:brazil}
\end{figure}

@@

@@

>>>>

\usepackage{natbib}
%\bibliographystyle{elsarticle-harv}

@@

A few years ago, \citep{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP), and {\it Weight and Balance Problem} (WBP). Several aspects were considered: item characteristics to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs); when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how total cargo weight is balanced; etc.

Recently, \citep{zhao2023} developed integer programming solution that address ACLPP and WBP separately, impacting payload optimization and aircraft centre of gravity (CG).
The authors argue that, with improvements in computer processing power, joint combinatorial optimization of ACPP and WBP is now feasible. They propose three integer linear programming models: a bi-objective optimization model (BOM), a combinatorial optimization model (COM), and an improved combinatorial optimization model (IOM). The models aim to determine the maximum loading capacity and lowest CG deviation from a specified target CG.

The study uses Gurobi to solve four scenarios with various conditional metrics for three models. Results show that the BOM has the fastest solution speed, but the CG deviation is the largest, making the results unacceptable in some cases. The COM has the longest solution time, making it difficult to tolerate in practice. The IOM offers the best optimization solution, despite taking a little longer to solve computationally than the BOM.

Using a Boeing 777F, four scenarios and 25 cases were tested, finding that the BOM's variable feasible domain is limited, the optimization result is insufficient, and the CG deviation is sometimes too large. The COM takes too long to solve, making it difficult to meet the needs of the actual scene. The IOM accelerates the solving process compared to the COM, but some scenarios still require significant computation time.

@@

\begin{table}[htp]

@@

\begin{table}[htp]
	\centering
	\caption{Air cargo transport: literature, sub-problems and features}  \label{tab:sa}
	\footnotesize
	\begin{tabular}{r|cccc|ccccc}
		\toprule
		Work & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\citep{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\citep{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\citep{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\citep{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\citep{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{Paquay2016,Paquay2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\citep{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\citep{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		\citep{zhao2023} & $\bigstar$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		\citep{MiguelOptimalAP} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		{\bf This article}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

\vspace{-2mm}

@@

\begin{table}[htp]
	\centering
	\caption{Aircraft parameters}  \label{tab:larger}
	\footnotesize
	\begin{tabular}{c | c c c c c c c c c}
		\toprule
		& \multicolumn{3}{c}{$Payload$: 75,000kg} & \multicolumn{3}{c}{$limit^{CG}_{long}$: $1.170m$} &
		\multicolumn{3}{c}{$limit^{CG}_{lat}$: $0.19m$} \\
		\midrule
		\multirow{2}{*}{{\boldm{$p_i$}}}  & $p_{17}$ & $p_{15}$ & $p_{13}$ & $p_{11}$ & $p_{9}$ & $p_{7}$ & $p_{5}$ & $p_{3}$ & $p_{1}$ \\
		& $p_{18}$ & $p_{16}$ & $p_{14}$ & $p_{12}$ & $p_{10}$ & $p_{8}$ & $p_{6}$ & $p_{4}$ & $p_{2}$ \\
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{long}$} ($m$)} & -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\
		& -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\			
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{lat}$} ($m$)}  & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 \\
		& -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 \\	
		\midrule
		{\boldm{$W_i$}} ($kg$)      &   4,500   &    4,500  &   4,500   &  4,500    & 4,500     & 4,500     & 4,500     & 3,000    & 3,000   \\
		{\boldm{$V_i$}} ($m^3$)   &   14.8   &   14.8   &  14.8    &  14.8    & 14.8     & 14.8     & 14.8     & 10.0    & 7.0 \\	
		\midrule	
		\textbf{Fuel cost}  & \multicolumn{9}{c}{ $c_d$ = US\$ $4.90/km$ }	 \\
		\midrule
		\textbf{Fuel consumption rate}  & \multicolumn{9}{c}{ $c_g = 5\%$} \\		
		\midrule
		\textbf{Maximum weight}  & \multicolumn{9}{c}{ $W_{max} = \sum_i W_i = 75,000kg$} \\
		\bottomrule
	\end{tabular}
	\normalsize 
\end{table}

@@

\begin{figure}[htp]
	\centering
	\includegraphics[scale=0.22]{lateral.png}
	\caption{Aircraft longitudinal cut, where red lines are pallets positions}
	\label{fig:lateral}
\end{figure}

@@

\begin{table}[htp]
	\centering
	\footnotesize
	\caption{Problem structure}  \label{tab:struc}
	\begin{tabular}{l l}
		\toprule
		\textbf{Notation}	& \textbf{Description} \\
		\midrule
		$L = \{ 0, 1, \ldots, K \}$ & Set of $K+1$\/ nodes of the tour, where node 0 is the base \\
		\midrule
		$\pi$               & A permutation between nodes 1, \ldots, $K$\\
		\midrule
		$S_K$               & Set of $K!$\/ permutations \\
		\midrule
		$\pi_k$             & The $k^{th}$\/ node of tour $\pi$, $1 \leq k \leq K$\\
		\midrule
		Tour $\pi$          & $\{0,\pi_1,\ldots, \pi_K, 0\}$ \\
		& For ease of notation, $\pi_0 = \pi_{K+1} = 0$\\
		\midrule
		$L_k$               & Set of remaining nodes of tour at node $k$, $0 \leq k \leq K$\\
		& By definition, $L_0 = L$\\
		\midrule
		$d(a,b)$            & Distance from node $a$\/ to node $b$, where $0 \leq a, b \leq K$ \\
		&   By definition, $d(a,a) = 0, \forall a$\\
		\midrule
		$C=\left[c_{a,b}\right]$    & Cost matrix of flights, where $c_{a,b} = c_d \times d(a,b)$\\
		\midrule
		$M = \{1, \ldots, m \}$  & Set of $m$\/ pallets in specific positions within the aircraft \\
		&  See Table \ref{tab:larger}, where $m=18$ \\
		\midrule
		$N_k = \{1, \ldots, n_k \}$ & Set of $n_k$\/ items available for loading at node $k$, $1 \leq j \leq n_k$, $0 \leq k \leq K$\\
		\midrule
		$N = \bigcup_{0 \leq k \leq K} N_k$ & Set of items in all nodes along a tour\\
		\midrule
		$Q_k = \{1, \ldots, m_k \}$ & Set of $m_k \leq m$\/ packed contents at node $k$, $1 \leq q \leq m_k$, $0 \leq k \leq K$ \\ 
		& By definition, $m_0 = 0$ and $Q_0 = \varnothing$\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[htp]
	\centering
	\footnotesize
	\caption{Decision variables and allocation graph}  \label{tab:graph}
	\begin{tabular}{l l}
		\toprule
		\textbf{Notation}	& \textbf{Description} \\
		\midrule
		$X_{ij}^{\pi_k}$\/ and $Y_{iq}^{\pi_k}$ & Binary variables, where $1 \leq i \leq m$, $1 \leq j \leq n_{\pi_k}$, $1 \leq q \leq m_{\pi_k}$\/ and $0 \leq k \leq K$\\
		\midrule
		$X_{ij}^{\pi_k} = 1$ & If item $j$\/ at node ${\pi_k}$\/ is assigned to pallet $i$, and 0 otherwise\\
		\midrule
		$Y_{iq}^{\pi_k} = 1$ & If packed content $q$\/ at node ${\pi_k}$\/ is assigned to pallet $i$, and 0 otherwise\\
		\midrule
		$T_i^{\pi_k} \in L_{\pi_k}$ &  Destination of items and packed contents assigned to pallet $i$\ at node ${\pi_k}$\\
		\midrule
		$G_{\pi_k}(V_{\pi_k}, E_{\pi_k})$ & Allocation graph at node ${\pi_k}$\\
		\midrule
		$V_{\pi_k} = M \cup N_{\pi_k} \cup Q_{\pi_k}$ & Allocation graph vertices at node ${\pi_k}$: pallets, items and packet contents\\
		\midrule
		$E_{N_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$, corresponding to shipped items\\
		\midrule
		$E_{Q_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$, corresponding to packed contents\\
		\midrule
		$E_{\pi_k} = E_{N_{\pi_k}} \cup E_{Q_{\pi_k}}$ & Allocation graph edges at node ${\pi_k}$\\
		\midrule
		$(i, j) \in E_{N_{\pi_k}}$ & If $X_{ij}^{\pi_k} = 1$, where $i$\/ is a pallet and $j$\/ is a item at node ${\pi_k}$ \\
		\midrule
		$(i, q) \in E_{Q_{\pi_k}}$ & If $Y_{iq}^{\pi_k} = 1$, where $i$\/ is a pallet and $q$\/ is a packed content at node ${\pi_k}$\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[htp]
	\centering
	\footnotesize
	\caption{Calculus functions} \label{tab:functions}
	\begin{tabular}{ c l }
		\toprule
		\textbf{Function} & \textbf{Description} \\
		\midrule
		(\ref{eq:scores}) & Total score of transported items throughout tour $\pi$\\
		\midrule
		(\ref{eq:tau})    & Longitudinal torque applied by loaded pallets at node ${\pi_k}$\\
		\midrule
		(\ref{eq:costs})  & Total cost of fuel on tour $\pi$\/ (distances and CG longitudinal deviations) \\
		\midrule
		(\ref{eq:departing}) & Set of not visited nodes at node ${\pi_k}$\\
		\midrule
		(\ref{eq:LatItem}) & Lateral torque at node ${\pi_k}$ (shipped items) \\
		\midrule
		(\ref{eq:LatPacked}) & Lateral torque at node ${\pi_k}$ (packed contents)\\
		\midrule
		(\ref{eq1}) & Objective function of ACLP+RPDP\\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

\begin{table}[htp]
	\centering
	\footnotesize
	\caption{Constraints} \label{tab:constraints}
	\begin{tabular}{ c l }
		\toprule
		\textbf{Constraint} & \textbf{Description} \\
		\midrule
		(\ref{eq:torqlong}, \ref{eq:torqlat}) & Longitudinal and lateral torques must be within aircraft limits \\
		\midrule
		(\ref{eq:app2}, \ref{eq:app3}) & Items allocated to each pallet cannot exceed its weight and volume limits\\
		\midrule
		(\ref{eq:app4}) & At most, each item is associated with a single pallet \\
		\midrule
		(\ref{eq:app5}) & Packed contents that have not yet reached their destination must remain on board\\
		\midrule
		(\ref{eq22}, \ref{eq23}) & Items allocated on the same pallet must have the same destinations\\
		\midrule
		(\ref{eq24}, \ref{eq25}) & If there is a packed content on the pallet, it must also have the same destination as other items\\

@@

\begin{algorithm}[htp]

@@

In the generation of test instances, we use two types of random selections:

\vspace{-2mm}

\begin{itemize}
	\setlength\itemsep{-0.3em}
	\item $RandomInt(i_1,i_2)$: randomly selects a integer number in $[i_1,i_2]$, where $i_1$ and $i_2$ are integer numbers;
	\item $Roulette(set)$ biased through $\phi$: selects an element from $set$, where the probability of each element is proportional to the value of a given function $\phi$\/ defined on $set$.
\end{itemize}

@@

	\begin{table}[H]
	\centering
	\caption{\textit{iRace} testing results}  \label{tab:irace}
	\footnotesize
	\begin{tabular}{cccc}
		\toprule

		\boldm{$surplus$} & \boldm{$threshold_1$} & \boldm{$threshold_2$} & \textbf{run time}\\
		& for $\eta_1$          & for $\eta_2$          & (min) \\
		\midrule
		1.2& 0.8621 & 1.0539 & 47 \\
		1.5& 0.9199 & 1.1399 & 59 \\
		2.0& 0.9617 & 1.5706 & 63 \\						
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

	\centering
	\caption{ \color{blue}Overall results}  \label{tab:mr}
	\footnotesize
	\begin{tabular}{cccc}	
	\toprule
	{\bf Method} & {\bf Best scenarios} & {\bf Worst scenarios} & {\bf Worst run times (min)}  \\
	\midrule
	NMO           & 4             & 5    & {\color{red}60}  \\
	ACO           & 2, 3          & 4, 5 & {\color{red}25}, {\color{red}61}  \\
	GRASP         & 1             & 4, 5 & {\color{red}28}, {\color{red}55}  \\
	TS            & -             & 5    & {\color{red}44}  \\
	\midrule
	{\it Gurobi}  & 1, 2, 3, 4    & 5    & {\color{red}did not solve}  \\
	{\it Shims}   & 1, 2, 3, 4, 5 & -    & {\bf 3.26} \\
	\bottomrule
	\end{tabular}
\normalsize

\end{table}

@@

	\begin{figure}[H]
		\centering
		\caption{\label{fig:brazil} {\it Shims/GA} best ACLP+RPDP tour \\
			\tiny{Solution for scenario 7, instance 1}}
		\fbox{\includegraphics[scale=0.30]{brazil.png}}
	\end{figure}

@@

It is important to remember that this is not the shortest tour, but the one with the best relationship with score, cost, and fuel consumption.

@@

{\color{purple}

	It is also important to highlight that three strategies were essential for us to obtain a solution in the instances created:

	\vspace{-2mm}

	\begin{itemize}
		\setlength\itemsep{-0.3em}
		\item[(1)] consider $K<m$\ so that each pallet received a predefined destination;
		\item[(2)] on each tour, calculate a node-by-node solution; and
		\item[(3)] at each intermediate node of the considered tour, relocate the packed contents, minimizing the torque on the aircraft.
	\end{itemize}

	\vspace{-2mm}

	Without adopting these three strategies, we were unable to carry out any experiments, even in smaller instances.
}
